---
title: "Bird Joins"
author: "Wylie Hampson"
date: "8/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(lubridate)
```

## Read in the csv files.

```{r}
bird_observations <- read_csv(here::here("data", "52_pp52_birds_1.csv"))
sites <- read_csv(here::here("data", "52_pp52_sites_1.csv"))
surveys <- read_csv(here::here("data", "52_pp52_surveys_1.csv"))
taxalist <- read_csv(here::here("data", "52_pp52_taxalist_1.csv"))
```

### Create a bird_subset for birds with species ID BHCO and RWBL and from sites LI-W and NU-C

```{r}
birds_subset <- bird_observations %>% 
  filter(species_id %in% c("BHCO", "RWBL")) %>% 
  filter(site_id %in% c("LI-W", "NU-C"))
```

### Use a left_join to include information from the sites and taxalists data frames.

```{r}
birds_left <- left_join(birds_subset, sites, by = "site_id")

birds_left <- left_join(birds_left, taxalist, by = "species_id")
```

### Use a full_join to add information from site to bird_subset

```{r}
birds_full <- full_join(birds_subset, sites, by = "site_id")
```

### Use an inner_join to join birds_subset and taxalist

```{r}
birds_inner <- inner_join(birds_subset, taxalist, by = "species_id")
```

### Try it! What happens in the above situation when you use left_join instead of inner_join?

```{r}
birds_left_test <- left_join(birds_subset, taxalist, by = "species_id")
```

## Task 2

rename the notes in bird_observations to bird_obs_notes

```{r}
birds <- bird_observations %>% 
  rename(bird_obs_notes = notes)
```

### Join the 4 datasets and then limit them to the survey_date, common_name, park_name, and bird_count columns.

```{r}
birds_joined <- left_join(birds, taxalist, by = "species_id")
birds_joined <- left_join(birds_joined, sites, by = "site_id")
birds_joined <- left_join(birds_joined, surveys, by = c("survey_id", "site_id"))

birds_joined <- birds_joined %>% 
  select(survey_date, common_name, park_name, bird_count)
```

### Add a new column to the above data frame called survey_month, then convert the month number to a factor

```{r}
birds_joined <- birds_joined %>% 
  mutate(survey_month = lubridate::month(survey_date)) %>% 
  mutate(survey_month = as.factor(survey_month)) %>% 
  dplyr::relocate(survey_month, .after = survey_date)
```

### Find the number of birds observed in each park by month

```{r}
bird_summary <- birds_joined %>% 
  group_by(park_name, survey_month) %>% 
  summarize(count = sum(bird_count, na.rm = TRUE)) %>% 
  filter(park_name %in% c("Lindo", "Orme", "Palomino", "Sonrisa"))

bird_summary
```








